Network Structure, Efficiency, and Performance in WikiProjects

### 2018-01-09 paper-content.tex
Revise coeditor networks, and empirical results.

Empirically, we find that both efficiency and performance are highly right-skewed,
with a small number of projects having values much higher than the average.
After log-transforming the values, both the efficiency and the performance have
a unimodal distribution with low skew (see Figure \ref{fig:eff-perf-hist}).

@@

The {\em coeditor network} of a WikiProject consists of nodes representing editors,
and edges connecting any editors who have edited the same article.
The edges are directed, with the direction representing the direction of
{\em plausible information flow};
an edge from Alice to Bob exists if Alice edited an article and then Bob edited the same article at
a later time.
Edges can exist in both directions e.g.,
if an article was edited first by Alice, then by Bob, and again by Alice.
We make the simplifying assumption of unit weight for all edges.
We focus on three structural properties:
degree, characteristic path length, and min-cut.
Degree and characteristic path length have been shown to correlate with
performance and efficiency in some social learning settings
\cite{golub_naive_2010,mason_propagation_2008,grim_scientific_2013},
while min-cut can be interpreted as a measure of decentralization,
a key property of Wikipedia.

@@

We also calculate the characteristic path length for each WikiProject coeditor network.
The {\em distance} from editor A to editor B is the length of the shortest path from A to B.
The {\em characteristic path length} is the mean distance between all editor pairs.
If no path exists between two editors, we exclude that pair from the mean.
For brevity, we will simply refer to this quantity as the {\em path length}.
The path length represents how quickly information can move through the network.
Networks with longer paths require more interactions for information to propagate through
the network,
which has been shown to reduce efficiency in some settings
\cite{mason_propagation_2008,barkoczi_social_2016}.

@@

The mean path and min-cut network measures are computationally intensive,
requiring distance and minimum $st$-cut calculations for each all node pairs.
For larger projects, these calculations are impractical.
When coeditor networks were large, we employed sampling to determine mean path length and mean
min-cut.
For mean path length, source nodes were sampled, and path length was calculated to all destination nodes
from each of these.
For min-cut, node pairs were sampled.
In both cases, stratification was used to ensure the same number of nodes were were sampled from each of
12 node degree quantiles.
We estimated the error due to sampling by determining true values for a medium-sized project,
and calculating error as a function of sample-size.
Sample sizes were chosen such that relative error was below 10\%.
Even with sampling, however, it was impractical to calculate these properties for the largest projects,
so we exclude them from the analysis.
To control for bias in project size, we include several size-related variables in our models.

\subsection{Regression Models}

@@

We model performance and efficiency in WikiProjects using ordinary least-squares linear regression.
Each WikiProject is taken as a single observation.
The models include each project's coeditor network properties as independent variables.
We also include several project-level variables to control for confounding factors.
The C-efficiency is included in the model for performance to control for the presence of articles
that are actively being improved.
Projects with lower efficiency will have more works-in-progress and could have an artificially low
performance without controlling for efficiency.
Some variables are log-transformed to reduce heteroscedasticity.

Our models are summarized in Table \ref{tab:model}.
Mean min-cut was found to be highly correlated with degree ($r=0.980$, $p<0.001$),
so we exclude min-cut to prevent collinearity (see Figure \ref{fig:degree-mincut}.
The high correlation between mean degree and min-cut implies that in most cases
the minimum $st$-cut is simply the either set of edges from $s$ or the set of edges to $t$.
The rarity of non-trivial min-cuts suggests that WikiProject coeditor networks have very few central
bottlenecks and are thus highly decentralized.
In-degree and out-degree skewness were also highly correlated, so each dependent variable was modeled twice:
once with in-degree skewness and once with out-degree skewness.

The regression results are consistent whether in-degree or out-degree skewness is used,
although the out-degree models are slightly more significant when modeling efficiency.
We also see that B-efficiency and C-efficiency have very similar models, but that A-efficiency behaves
differently in its dependence on degree skewness and connectivity.
The different behavior of A-efficiency is likely explained by the observation that the A-Class quality is
infrequently used in practice, meaning that the quality level is usually achieved when an article is rated
as a good or featured article, which involves a different consensus process than lower ratings.

>>>>

@@

The {\em coeditor network} of a WikiProject consists of nodes representing editors,
and edges connecting any editors who have edited the same article.
The edges are directed, with the direction representing the direction of
{\em plausible information flow};
an edge from Alice to Bob exists if Alice edited an article and then Bob edited the same article at
a later time.
Edges can exist in both directions e.g.,
if an article was edited first by Alice, then by Bob, and again by Alice.
We make the simplifying assumption of unit weight for all edges.
We focus on three structural properties:
degree, characteristic path length, and min-cut.
Degree and characteristic path length have been shown to correlate with
performance and efficiency in some social learning settings
\cite{golub_naive_2010,mason_propagation_2008,grim_scientific_2013},
while min-cut can be interpreted as a measure of decentralization,
common feature of peer-produced communities such as Wikipedia
\cite{benkler_wealth_2006}.

@@

We also calculate the characteristic path length for each WikiProject coeditor network.
The {\em distance} from node $s$ to node $t$ is the distance of the shortest path
from $s$ to $t$.
The {\em characteristic path length} (or just {\em path length} for brevity)
is the mean distance between all editor pairs,
excluding unconnected pairs.
To account for unconnected pairs, we also measure the {\em connected fraction}
of all possible edge pairs.
The path length represents how quickly information can move through the network.
Networks with longer paths require more interactions for information to propagate through
the network,
which has been shown to reduce efficiency in some settings
\cite{mason_propagation_2008,barkoczi_social_2016}.

@@

The mean path and min-cut network measures are computationally intensive,
requiring distance and minimum $st$-cut calculations for each all node pairs.
For larger projects, these calculations are impractical.
When coeditor networks were large, we employed sampling to determine mean path length and mean
min-cut.
For mean path length, source nodes were sampled, and path length was calculated to all destination nodes
from each of these.
For min-cut, node pairs were sampled.
In both cases, stratification was used to ensure the same number of nodes were were sampled from each of
12 node degree quantiles.
We estimated the error due to sampling by determining true values for a medium-sized project,
and calculating error as a function of sample-size.
Sample sizes were chosen such that relative error was below 10\%.
Even with sampling, however, it was impractical to calculate these properties for the largest projects,
so we exclude the 183 largest projects from the analysis.
To control for bias in project size, we include several size-related variables in our models.

\subsection{Empirical Results}

We find that both efficiency and performance are highly right-skewed,
with a small number of projects having values much higher than the average.
After log-transforming the values, both the efficiency and the performance have
a unimodal distribution with low skew (see Figure \ref{fig:eff-perf-hist}).

We also find that mean min-cut is highly correlated with degree ($r=0.980$, $p<0.001$),
so we exclude min-cut from regression models to prevent collinearity (see Figure \ref{fig:degree-mincut}).
The high correlation between mean degree and min-cut implies that in most cases
the minimum $st$-cut is simply the either set of edges from $s$ or the set of edges to $t$.
The rarity of non-trivial min-cuts suggests that WikiProject coeditor networks have very few central
bottlenecks and are thus highly decentralized.

@@

To study the relationship between network structure, efficiency, and performance,
we model the performance and efficiency of WikiProjects using ordinary least-squares linear regression.
Each WikiProject is taken as a single observation.
The models include each project's coeditor network properties as independent variables.
We also include several project-level variables to control for confounding factors,
described below.
\begin{description}
\item[C-efficiency] Quantifies how quickly a WikiProject improves low-quality articles.
\item[Connected fraction] Fraction of coeditor pairs connected by a path.
\item[Mean editors/article] Mean number of editors collaborating on each article in a WikiProject.
\item[Article count] Total number of articles in the WikiProject.
\item[Editor count] Total number of editors working on articles within a WikiProject.
\item[Revision count] Total number of revisions to articles in a WikiProject.
\item[First assessment] Timesamp of first assessment; a measure of how long a WikiProject has been active.
\item[Mean article age] Mean age of articles within a WikiProject.
\item[Mean similarity] Mean Jaccard similarity (by article) with other WikiProjects; a measure of topical complexity.
\end{description}

Our models are summarized in Table \ref{tab:model}.
Min-cut is excluded from all models to avoid collinearity,
as it is highly correlated with degree.
In-degree and out-degree skewness were also highly correlated,
so we only include out-degree skewness in the models.
Some variables are log-transformed.
To test the robustness of our results,
we also computed models using cube root instead of logarithmic transformations,
as well as using only top-importance and high-importance articles.
These robustness check yielded qualitatively similar results for all variables,
except for degree-skewness, which had an inconsistent sign across models.

We see that B-efficiency and C-efficiency have very similar models, but that A-efficiency behaves
differently in its dependence on degree skewness and connectivity.
The different behavior of A-efficiency is likely explained by the observation that the A-Class quality is
infrequently used in practice, meaning that the quality level is usually achieved when an article is rated
as a good or featured article, which involves a different consensus process than lower ratings.
